---
title: "TPL Themes"
author: "Connor Rothschild"
date: "Last updated 7/15/19"
output: 
  rmarkdown::html_vignette:
    fig_width: 7
    fig_height: 4
vignette: >
  %\VignetteIndexEntry{my-vignette}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)

knitr::opts_knit$set(
  root.dir = normalizePath("../")
)
```

```{r, warning = FALSE, message = FALSE}
if (!require('devtools')) install.packages('devtools')
config <- utilsR::read_yaml("config.yaml")
library(ggplot2)
#library(tpltheme)
devtools::load_all()
```

## Plotting

### Overview

This package creates a standardized formats for plots to be used in reports created by the Texas Policy Lab. It primarily relies on `set_tpl_theme`, which allows the user to specify whether the plot theme should align with a standard plot (`style = "print"`), or one specially created for plotting geographical data (`style = "Texas"`).

### Fonts

The user is able to specify whether they want to use *Lato* or *Adobe Caslon Pro* in their figures.

To ensure that these fonts are installed and registered, use `font_test()`. If fonts are not properly installed, install both fonts online and then run `font_install()`.

```{r, warning = FALSE}
font_test()
font_install()
```

### Usage

Load `library(tpltheme)` **after** `library(ggplot2)` and/or `library(tidyverse)`.

Here are some examples of sample TPL plots with different specifications for `style` and `font`.

```{r}
#library(tidyverse)
#library(tpltheme)
set_tpl_theme(style = "print", font = "adobe")

ggplot(iris, aes(x=Species, y=Sepal.Width, fill=Species)) +
    geom_bar(stat="summary", fun.y="mean", show.legend = FALSE) +
    labs(x="Species", y="Mean Sepal Width (cm)", fill="Species", title="Iris Dataset")
```
```{r}
set_tpl_theme(style = "print", font = "lato")

ggplot(iris, aes(x=jitter(Sepal.Width), y=jitter(Sepal.Length), col=Species, size = Petal.Length)) +
    geom_point() +
    labs(x="Sepal Width (cm)", y="Sepal Length (cm)", col="Species", size = "Petal Length", title="Iris Dataset")
```

By specifying `style = "Texas"` within `set_tpl_theme`, the user may also create Texas-specific plots.

```{r, warning = FALSE, message = FALSE}
tx_vac <- readr::read_csv("https://raw.githubusercontent.com/connorrothschild/tpltheme/master/tx_vac_example.csv")
set_tpl_theme(style = "Texas", font = "adobe")
ggplot2::ggplot(data = tx_vac, mapping = ggplot2::aes(x = long, y = lat, group = group, fill = avgvac*100)) +
  ggplot2::coord_fixed(1.3) +
  scale_fill_continuous(limits = c(78.3,100)) +
  ggplot2::geom_polygon(color = "black") +
  labs(title = "Texas Vaccination Rate by County",
       subtitle = "Among Kindergarteners",
       fill = "Percent\nVaccinated",
       caption = "Source: Texas DSHS")
```

And it also works for categorical variables:

```{r}
tx_vac %>% 
  dplyr::mutate(cat = factor(dplyr::case_when(avgvac*100 > 99 ~ "Great",
                         avgvac*100 > 90 ~ "Average",
                         avgvac*100 < 90 ~ "Bad"))) %>% 
  ggplot2::ggplot(mapping = ggplot2::aes(x = long, y = lat, group = group, fill = cat)) +
  ggplot2::coord_fixed(1.3) +
  ggplot2::geom_polygon(color = "black") +
  labs(title = "Texas Vaccination Rate by County",
       subtitle = "Among Kindergarteners",
       fill = "Vaccination Rating",
       caption = "Source: Texas DSHS")
```

If the number of colors exceeds the number of colors in the TPL palette (9), the function `tpl_color_pal()` will drop the TPL color palette and return the greatest number of unique colors possible within the RColorBrewer's "Paired" palette (for more information on the use of RColorBrewer palettes, see [this chapter](https://bookdown.org/rdpeng/exdata/plotting-and-color-in-r.html#using-the-rcolorbrewer-palettes)).

```{r, warning = FALSE, message = FALSE}
tx_vac %>% 
  ggplot2::ggplot(mapping = ggplot2::aes(x = long, y = lat, group = group, fill = subregion)) +
  ggplot2::coord_fixed(1.3) +
  ggplot2::geom_polygon(color = "black", show.legend = FALSE) +
  labs(title = "Texas Counties")
# default to print afterwards
set_tpl_theme(style = "print")
```

### TPL Branding

#### Logo

The user also has the option to include the TPL logo in single plots. This may be preferred for those reports being made especially public, or to serve as a pseudo-watermark in proprietary plots.

``` {r, message = FALSE, results ='hide'}
library(grid)
library(gridExtra)
plot <- ggplot(iris, aes(x=jitter(Sepal.Width), y=jitter(Sepal.Length), col=Species, size = Petal.Length)) +
    geom_point() +
    labs(x="Sepal Width (cm)", y="Sepal Length (cm)", col="Species", size = "Petal Length", title="Iris Dataset")
 
add_tpl_logo(plot)
```

The user can specify the `position` of the logo as well as it's `size`:

```{r, message = FALSE, results = 'hide'}
add_tpl_logo(tpl_plot_test(type = "barplot"), size = "regular", position = "bottomright")
```

Finally, the user is able to move the logo horizontally by specifying `align`. Positive values will shift the logo rightward while negative values shift it leftward:

```{r, message = FALSE, results = 'hide'}
add_tpl_logo(tpl_plot_test(type = "barplot"), size = "small", position = "bottomright", align = -2)
```

#### Logo Text

There may be some instances when an all-out logo is not warranted or preferred. If that is the case and the user would still like to watermark their figures, they can use the function `add_tpl_logo_text()` to add text to an existing plot object:

```{r}
plot <- ggplot(iris, aes(x=jitter(Sepal.Width), y=jitter(Sepal.Length), col=Species, size = Petal.Length)) +
    geom_point() +
    labs(x="Sepal Width (cm)", y="Sepal Length (cm)", col="Species", size = "Petal Length", title="Iris Dataset")
    
add_tpl_logo_text(plot)
```

The user may also need to specify `align`, which moves the plot horizontally across the bottom of the page.

```{r}
plot <- ggplot(iris, aes(x=Species, y=Sepal.Width, fill=Species)) +
    geom_boxplot(show.legend = FALSE) +
    labs(x="Species", y="Sepal Width (cm)", fill="Species", title="Iris Dataset", subtitle ="When specifying align = 1.5")
    
add_tpl_logo_text(plot, align = 1.5)    
```

### Additional Functions

#### Drop Axes

In the event that the user wishes to drop an axis, they may do so with `drop_axis()`. The function may drop any combination of axes depending on the user's input (`drop = "x"`, `drop = "y"`, `drop = "both"`, `drop = "neither"`).

Unlike `add_tpl_logo()`, `drop_axis()` should be *added* to an existing plot object:

```{r}
ggplot(iris, aes(x=jitter(Sepal.Width), y=jitter(Sepal.Length), col=Species, size = Petal.Length)) +
    geom_point() +
    labs(x="Sepal Width (cm)", y="Sepal Length (cm)", col="Species", size = "Petal Length", title="Iris Dataset") +
    drop_axis(axis = "y")
```

#### Color Palettes

The function `view_palette` plots base color palettes included in `tplthemes`. All TPL color palettes are led by the notation
`palette_tpl_*` and therefore can be easily autocompleted within RStudio.

```{r}
view_palette(palette = palette_tpl_diverging)
view_palette(palette = palette_tpl_main)
```

#### Restore Defaults

By calling `undo_tpl_theme`, you are able to remove TPL-specific theme settings and restores to ggplot defaults.

```{r}
undo_tpl_theme()
ggplot(iris, aes(x=jitter(Sepal.Width), y=jitter(Sepal.Length), col=Species, size = Petal.Length)) +
    geom_point() +
    labs(x="Sepal Width (cm)", y="Sepal Length (cm)", col="Species", size = "Petal Length", title="Iris Dataset")
```

## Reporting

* `read_word`: Reads word into Rmarkdown, such that word documents can be edited and read into the main Rmarkdown file for creating reports. 
* `read_word_table`: Reads a table from word into Rmarkdown, such that tables in word can be edited and then imported into Rmarkdown.
